|

### 2020-05-19 paper/Manuscript.Rmd
Last update of 19-5

<!--Barbara: Maybe write here something about writing manuscripts on a line-by-line basis, so that version control can keep track of changes more accurately than if a whole paragraph is identified as altered?-->

## Open science as opportunity

<!-- LB:  this section comes at a weird place in the flow of the paper. 
Barbara: agreed! I would remove this altogether.-->
Sometimes, open science principles are discussed as a kind of "antidote" to scientific misconduct. Yet these principles also present an opportunity to conduct more reliable research, with fewer mistakes, more convenient collaboration, more effective internal and external collaboration, and the potential for crowd-sourced error correction.

@@

TOP encourages comprehensive citation of literature, data, materials, methods, and software. In principle, researchers can meet this requirement by simply citing every reference used. Unfortunately, citation of data and software is less commonplace than citation of literature and materials. Crediting these resources is important, because it incentivizes data sharing and the development of open-source software, and supports the open science efforts of others.

<!--Various initiatives are underway to innovate how scientific output is measured and valued. https://www.nwo.nl/beleid/statement+waarderen+en+belonen+van+wetenschappers --> 
To facilitate citing datasets, researchers sometimes publish *data papers*; documents that detail the procedure, sample, and codebook. Specialized journals, such as the [*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/), aid in the publication of these data papers. For smaller projects, researchers sometimes publish their data in an online repository - like GitHub - along with a text file with the preferred citation for the data (which can be a substantive paper), and the license that applies to the data, such as Creative Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or [BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt, one can always contact the data creators and ask what the preferred citation is.

<!-- AL: Some general things about software citation that could be good to mention here: Tool or software papers are also common, often as short papers, sometimes called "Application Notes". The Journal of Open Source Software (JOSS) is a way of turning software directly into a publication, without the need of writing an actual paper about it (the review is about the code). https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.--> 
When using R, software (R packages) can be conveniently cited by calling `citation("packagename")`. This returns an APA-style reference, and a BibTeX entry. However, users of the WORCS project template for RStudio do not need to manually cite R-packages. WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.

>>>>

@@

TOP encourages comprehensive citation of literature, data, materials, methods,
and software. In principle,
researchers can meet this requirement by simply
citing every reference used. Unfortunately, citation of data and software is
less commonplace than citation of literature and materials. Crediting these
resources is important, because it incentivizes data sharing and the development
of open-source software, supports the open science efforts of others, and helps
researchers receive credit for *all* of their research output, in line with
[DORA](https://sfdora.org/).

To facilitate citing datasets, researchers sometimes publish *data papers*;
documents that detail the procedure, sample, and codebook. Specialized journals,
such as the
[*Journal of Open Psychology Data*](https://openpsychologydata.metajnl.com/),
aid in the publication of these data papers. For smaller projects, researchers
sometimes publish their data in an online repository - like 'GitHub' - along
with a text file with the preferred citation for the data (which can be a
substantive paper), and the license that applies to the data, such as Creative
Commons [BY-SA](https://creativecommons.org/licenses/by-sa/4.0/) or
[BY-NC-SA](https://creativecommons.org/licenses/by-nc-sa/4.0/). When in doubt,
one can always contact the data creators and ask what the preferred citation is.

<!-- AL: Some general things about software citation that could be good to mention here: Tool or software papers are also common, often as short papers, sometimes called "Application Notes". The Journal of Open Source Software (JOSS) is a way of turning software directly into a publication, without the need of writing an actual paper about it (the review is about the code). https://citation-file-format.github.io/ is a file format to include software citation information into source code repositories.--> 
When using R, software (R packages) can be conveniently cited by calling
`citation("packagename")`. This returns an APA-style reference, and a BibTeX
entry. In addition to the manual citation of important packages, the `worcs`
project template for RStudio offers a comprehensive record of all dependencies. WORCS generates a template manuscript in APA6 style, based on the `papaja` package [@austPapajaPrepareReproducible2020], which automatically cites all packages used.
